Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update methods section for spelling mistakes

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment \cite{liu2005using,treviso2017sentence}. To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary. For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}. Since pre-trained word2vec word vectors provides the best results, in our experiment, we utilized pre-trained word2vec word vectors for words. Trained word vectors were also utilized when punctuation marks are not available in the pre-trained word2vec word vectors. Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.

@@

\textbf{Multilayer Perceptrons}: MLP is a feed forward artificial neural network model which maps set of input onto a set of appropriate output. This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase \cite{srivastava2014dropout}. Dropout is also applied in the fully connected layers in BRNN and CRNN models. For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.

@@

A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_c.n_e$ a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After the convolutional layer, $m \times n_f$ are fed into bidirectional recurrent neural networks with 200 hidden units. Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not. We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

>>>>

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment \cite{liu2005using,treviso2017sentence}. To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary. For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}. Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words. Trained word vectors were also utilized when punctuation marks are not available in the pre-trained word2vec word vectors. Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.

@@

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network model which maps set of input onto a set of appropriate output. This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase \cite{srivastava2014dropout}. Dropout is also applied in the fully connected layers in BRNN and CRNN models. For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors.

@@

A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After the convolutional layer, $m \times n_f$ are fed into bidirectional recurrent neural networks with 200 hidden units. Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not. We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.
